Arrows for Parallel Computation

### 2019-12-27 paper/content/arxiv/ms.tex
it compiles

% More experiences with seamless porting of parallel PArrows-based programs across the backends are welcome.
% Of course, we are ourselves working on expanding
In ongoing work we are expanding 
both our skeleton library and the number of skeleton-based parallel programs that use our DSL.
% to be portable across flavours of parallel Haskells.
It would also be interesting to see a hybrid of PArrows and Accelerate \cite{McDonell:2015:TRC:2887747.2804313}.
Ports of our approach to other languages such as Frege, Eta, or Java directly are at an early development stage.
	%\pagebreak
        %\bibliographystyle{jfp}
	%\bibliography{references,main}
        \providecommand{\noopsort}[1]{}
\begin{thebibliography}{}

\bibitem[\protect\citename{Acar {\em et~al.}\relax,
  }2000]{Acar:2000:DLW:341800.341801}
Acar, Umut~A., Blelloch, Guy~E., \& Blumofe, Robert~D. (2000).
\newblock The data locality of work stealing.
\newblock {\em Pages  1--12 of:} {\em Proceedings of the 12\xth Annual ACM
  Symposium on Parallel Algorithms and Architectures}.
\newblock SPAA '00.
\newblock ACM.

\bibitem[\protect\citename{Achten {\em et~al.}\relax, }2007]{achten2007arrow}
Achten, Peter, van Eekelen, Marko, de~Mol, Maarten, \& Plasmeijer, Rinus.
  (2007).
\newblock An arrow based semantics for interactive applications.
\newblock  {\em Draft Proceedings of the Symposium on Trends in Functional
  Programming}.
\newblock TFP '07.

\bibitem[\protect\citename{Achten {\em et~al.}\relax, }2004]{achten2004arrows}
Achten, PM, van Eekelen, Marko~CJD, Plasmeijer, MJ, \& Weelden, A~van. (2004).
\newblock {\em Arrows for generic graphical editor components}.
\newblock Tech. rept. Nijmegen Institute for Computing and Information
  Sciences, Faculty of Science, University of Nijmegen, The Netherlands.
\newblock Technical Report NIII-R0416.

\bibitem[\protect\citename{Alimarine {\em et~al.}\relax,
  }2005]{Alimarine:2005:BAA:1088348.1088357}
Alimarine, Artem, Smetsers, Sjaak, van Weelden, Arjen, van Eekelen, Marko, \&
  Plasmeijer, Rinus. (2005).
\newblock There and back again: Arrows for invertible programming.
\newblock {\em Pages  86--97 of:} {\em Proceedings of the {2005} {ACM}
  {SIGPLAN} Workshop on {Haskell}}.
\newblock Haskell '05.
\newblock ACM.

\bibitem[\protect\citename{Aljabri {\em et~al.}\relax,
  }2014]{Aljabri:2013:DIG:2620678.2620682}
Aljabri, Malak, Loidl, Hans-Wolfgang, \& Trinder, Phil~W. (2014).
\newblock The design and implementation of {GUMSMP}: A multilevel parallel
  haskell implementation.
\newblock {\em Pages  37:37--37:48 of:} {\em Proceedings of the 25\xth
  Symposium on Implementation and Application of Functional Languages}.
\newblock IFL '13.
\newblock ACM.

\bibitem[\protect\citename{Aljabri {\em et~al.}\relax, }2015]{Aljabri2015}
Aljabri, Malak, Loidl, Hans-Wolfgang, \& Trinder, Phil. (2015).
\newblock {\em Balancing shared and distributed heaps on {NUMA} architectures}.
\newblock Springer.
\newblock Pages  1--17.

\bibitem[\protect\citename{Alt \& Gorlatch, }2003]{AlGo03a}
Alt, Martin, \& Gorlatch, Sergei. (2003).
\newblock {F}uture-{B}ased {RMI}: Optimizing compositions of remote method
  calls on the {G}rid.
\newblock {\em Pages  682--693 of:} Kosch, Harald, B{\"o}sz{\"o}rm{\'e}nyi,
  L{\'a}szl{\'o}, \& Hellwagner, Hermann (eds), {\em Euro-Par~2003}.
\newblock LNCS 2790.
\newblock Springer-Verlag.

\bibitem[\protect\citename{Asada, }2010]{Asada:2010:ASM:1863597.1863607}
Asada, Kazuyuki. (2010).
\newblock Arrows are strong monads.
\newblock {\em Pages  33--42 of:} {\em Proceedings of the Third {ACM SIGPLAN}
  Workshop on Mathematically Structured Functional Programming}.
\newblock MSFP '10.
\newblock New York, NY, USA: ACM.

\bibitem[\protect\citename{Aswad {\em et~al.}\relax, }2009]{aswad2009low}
Aswad, Mustafa, Trinder, Phil, Al~Zain, Abdallah, Michaelson, Greg, \&
  Berthold, Jost. (2009).
\newblock Low pain vs no pain multi-core {Haskells}.
\newblock {\em Pages  49--64 of:} {\em Trends in Functional Programming}.

\bibitem[\protect\citename{Atkey, }2011]{ATKEY201119}
Atkey, Robert. (2011).
\newblock What is a categorical model of arrows?
\newblock {\em Electronic notes in theoretical computer science}, {\bf 229}(5),
  19--37.

\bibitem[\protect\citename{Berthold, }2008]{JostThesis}
Berthold, Jost. (2008).
\newblock {\em Explicit and implicit parallel functional programming ---
  concepts and implementation}.
\newblock Ph.D. thesis, Philipps-Universit\"{a}t Marburg.

\bibitem[\protect\citename{Berthold \& Loogen, }2006]{Eden:PARCO05}
Berthold, Jost, \& Loogen, Rita. (2006).
\newblock Skeletons for recursively unfolding process topologies.
\newblock  Joubert, Gerhard~R., Nagel, Wolfgang~E., Peters, Frans~J., Plata,
  Oscar~G., Tirado, P., \& Zapata, Emilio~L. (eds), {\em Parallel Computing:
  Current \& Future Issues of High-End Computing, ParCo 2005, Malaga, Spain}.
\newblock NIC Series 33.
\newblock Central Institute for Applied Mathematics, J{\"u}lich, Germany.

\bibitem[\protect\citename{Berthold \& Loogen, }2007]{Berthold2007a}
Berthold, Jost, \& Loogen, Rita. (2007).
\newblock {Visualizing Parallel Functional Program Executions: Case Studies
  with the Eden Trace Viewer}.
\newblock  {\em ParCo '07. Parallel Computing: Architectures, Algorithms and
  Applications}.
\newblock IOS Press.

\bibitem[\protect\citename{Berthold {\em et~al.}\relax, }2008]{PADL08HMWS}
Berthold, Jost, Dieterle, Mischa, Loogen, Rita, \& Priebe, Steffen. (2008).
\newblock Hierarchical master-worker skeletons.
\newblock  Warren, David~S., \& Hudak, Paul (eds), {\em Practical Aspects of
  Declarative Languages (PADL'08)}.
\newblock LNCS 4902.
\newblock San Francisco (CA), USA: Springer-Verlag.

\bibitem[\protect\citename{Berthold {\em et~al.}\relax, }2009a]{arcs-dc}
Berthold, Jost, Dieterle, Mischa, Lobachev, Oleg, \& Loogen, Rita. (2009a).
\newblock {Distributed Memory Programming on Many-Cores -- A Case Study Using
  {Eden} Divide-\&-Conquer Skeletons}.
\newblock {\em Pages  47--55 of:} Gro\ss{}pitsch, K.-E., Henkersdorf, A.,
  Uhrig, S., Ungerer, T., \& H\"{a}hner, J. (eds), {\em Workshop on Many-Cores
  at ARCS '09 -- 22\nd International Conference on Architecture of Computing
  Systems 2009}.
\newblock VDE-Verlag.

\bibitem[\protect\citename{Berthold {\em et~al.}\relax,
  }2009b]{Berthold2009-mr}
Berthold, Jost, Dieterle, Mischa, \& Loogen, Rita. (2009b).
\newblock Implementing parallel {Google} map-reduce in {Eden}.
\newblock {\em Pages  990--1002 of:} Sips, Henk, Epema, Dick, \& Lin, Hai-Xiang
  (eds), {\em Euro-Par 2009 Parallel Processing}.
\newblock LNCS 5704.
\newblock Springer Berlin Heidelberg.

\bibitem[\protect\citename{Berthold {\em et~al.}\relax,
  }2009c]{Berthold2009-fft}
Berthold, Jost, Dieterle, Mischa, Lobachev, Oleg, \& Loogen, Rita. (2009c).
\newblock {\em Parallel {FFT} with {Eden} skeletons}.
\newblock PaCT '09.
\newblock Springer.
\newblock Pages  73--83.

\bibitem[\protect\citename{Berthold {\em et~al.}\relax,
  }2016]{berthold_loidl_hammond_2016}
Berthold, Jost, Loidl, Hans-Wolfgang, \& Hammond, Kevin. (2016).
\newblock {PAEAN}: Portable and scalable runtime support for parallel {Haskell}
  dialects.
\newblock {\em Journal of functional programming}, {\bf 26}.

\bibitem[\protect\citename{Bischof \& Gorlatch, }2002]{Bischof2002}
Bischof, Holger, \& Gorlatch, Sergei. (2002).
\newblock {\em Double-scan: Introducing and implementing a new data-parallel
  skeleton}.
\newblock Euro-Par '02.
\newblock Springer.
\newblock Pages  640--647.

\bibitem[\protect\citename{Blumofe \& Leiserson,
  }1999]{Blumofe:1999:SMC:324133.324234}
Blumofe, Robert~D., \& Leiserson, Charles~E. (1999).
\newblock Scheduling multithreaded computations by work stealing.
\newblock {\em J. acm}, {\bf 46}(5), 720--748.

\bibitem[\protect\citename{Botorog \& Kuchen, }1996]{botorog1996efficient}
Botorog, G.~H., \& Kuchen, H. (1996).
\newblock {\em {Euro-Par'96 Parallel Processing}}.
\newblock LNCS 1123.
\newblock Springer-Verlag.
\newblock Chap. Efficient parallel programming with algorithmic skeletons,
  pages  718--731.

\bibitem[\protect\citename{Brown \& Hammond, }2010]{brown2010ever}
Brown, C., \& Hammond, K. (2010).
\newblock Ever-decreasing circles: a skeleton for parallel orbit calculations
  in {Eden}.
\newblock  {\em Draft Proceedings of the Symposium on Trends in Functional
  Programming}.
\newblock TFP '10.

\bibitem[\protect\citename{Buono {\em et~al.}\relax, }2010]{BUONO20102095}
Buono, D., Danelutto, M., \& Lametti, S. (2010).
\newblock Map, reduce and mapreduce, the skeleton way.
\newblock {\em Procedia computer science}, {\bf 1}(1), 2095--2103.
\newblock ICCS 2010.

\bibitem[\protect\citename{{C}hakravarty {\em et~al.}\relax,
  }2007]{Chakravarty2007}
{C}hakravarty, Manuel M.~T., {L}eshchinskiy, Roman, {P}eyton {J}ones, Simon~L.,
  {K}eller, Gabriele, \& {M}arlow, Simon. (2007).
\newblock {Data Parallel Haskell}: a status report.
\newblock {\em Pages  10--18 of:} {\em DAMP '07}.
\newblock ACM Press.

\bibitem[\protect\citename{Chakravarty {\em et~al.}\relax,
  }2011]{Chakravarty:2011:AHA:1926354.1926358}
Chakravarty, Manuel~M.T., Keller, Gabriele, Lee, Sean, McDonell, Trevor~L., \&
  Grover, Vinod. (2011).
\newblock Accelerating {Haskell} array codes with multicore {GPUs}.
\newblock {\em Pages  3--14 of:} {\em Proceedings of the 6\xth Workshop on
  Declarative Aspects of Multicore Programming}.
\newblock DAMP '11.
\newblock ACM.

\bibitem[\protect\citename{Chase \& Lev, }2005]{Chase:2005:DCW:1073970.1073974}
Chase, David, \& Lev, Yossi. (2005).
\newblock Dynamic circular work-stealing deque.
\newblock {\em Pages  21--28 of:} {\em Proceedings of the 17\xth Annual ACM
  Symposium on Parallelism in Algorithms and Architectures}.
\newblock SPAA '05.
\newblock ACM.

\bibitem[\protect\citename{Clifton-Everest {\em et~al.}\relax, }2014]{CMCK14}
Clifton-Everest, Robert, McDonell, Trevor~L, Chakravarty, Manuel M~T, \&
  Keller, Gabriele. (2014).
\newblock {Embedding Foreign Code}.
\newblock  {\em PADL '14: The 16th International Symposium on Practical Aspects
  of Declarative Languages}.
\newblock LNCS.
\newblock Springer-Verlag.

\bibitem[\protect\citename{{C}ole, }1989]{Cole1989}
{C}ole, M.~I. (1989).
\newblock Algorithmic skeletons: Structured management of parallel computation.
\newblock  {\em Research Monographs in Parallel and Distributed Computing}.
\newblock Pitman.

\bibitem[\protect\citename{Czaplicki \& Chong,
  }2013]{Czaplicki:2013:AFR:2499370.2462161}
Czaplicki, Evan, \& Chong, Stephen. (2013).
\newblock Asynchronous functional reactive programming for guis.
\newblock {\em {SIGPLAN} not.}, {\bf 48}(6), 411--422.

\bibitem[\protect\citename{Dagand {\em et~al.}\relax,
  }2009]{Dagand:2009:ORD:1481861.1481870}
Dagand, Pierre-\'{E}variste, Kosti\'{c}, Dejan, \& Kuncak, Viktor. (2009).
\newblock Opis: Reliable distributed systems in {OCaml}.
\newblock {\em Pages  65--78 of:} {\em Proceedings of the 4\xth International
  Workshop on Types in Language Design and Implementation}.
\newblock TLDI '09.
\newblock ACM.

\bibitem[\protect\citename{Danelutto {\em et~al.}\relax,
  }1992]{DANELUTTO1992205}
Danelutto, Marco, Meglio, Roberto~Di, Orlando, Salvatore, Pelagatti, Susanna,
  \& Vanneschi, Marco. (1992).
\newblock A methodology for the development and the support of massively
  parallel programs.
\newblock {\em Future generation computer systems}, {\bf 8}(1), 205--220.

\bibitem[\protect\citename{Darlington {\em et~al.}\relax,
  }1993]{darlington1993parallel}
Darlington, J., Field, AJ, Harrison, PG, Kelly, PHJ, Sharp, DWN, Wu, Q., \&
  While, RL. (1993).
\newblock Parallel programming using skeleton functions.
\newblock {\em Pages  146--160 of:} {\em Parallel architectures and languages
  Europe}.
\newblock Springer-Verlag.

\bibitem[\protect\citename{Dastgeer {\em et~al.}\relax,
  }2011]{Dastgeer:2011:ASM:1984693.1984697}
Dastgeer, Usman, Enmyren, Johan, \& Kessler, Christoph~W. (2011).
\newblock Auto-tuning {SkePU}: A multi-backend skeleton programming framework
  for multi-{GPU} systems.
\newblock {\em Pages  25--32 of:} {\em Proceedings of the 4\xth International
  Workshop on Multicore Software Engineering}.
\newblock IWMSE '11.
\newblock ACM.

\bibitem[\protect\citename{de~la Encina {\em et~al.}\relax,
  }2011]{delaEncina2011}
de~la Encina, Alberto, Hidalgo-Herrero, Mercedes, Rabanal, Pablo, \& Rubio,
  Fernando. (2011).
\newblock {\em A parallel skeleton for genetic algorithms}.
\newblock IWANN '11.
\newblock Springer.
\newblock Pages  388--395.

\bibitem[\protect\citename{Dean \& Ghemawat,
  }2008]{Dean:2008:MSD:1327452.1327492}
Dean, Jeffrey, \& Ghemawat, Sanjay. (2008).
\newblock {MapReduce}: simplified data processing on large clusters.
\newblock {\em Communications of the acm}, {\bf 51}, 107--113.

\bibitem[\protect\citename{Dean \& Ghemawat,
  }2010]{Dean:2010:MFD:1629175.1629198}
Dean, Jeffrey, \& Ghemawat, Sanjay. (2010).
\newblock {MapReduce}: a flexible data processing tool.
\newblock {\em Communications of the acm}, {\bf 53}, 72--77.

\bibitem[\protect\citename{Dieterle {\em et~al.}\relax, }2010a]{Dieterle2010}
Dieterle, M., Horstmeyer, T., \& Loogen, R. (2010a).
\newblock Skeleton composition using remote data.
\newblock {\em Pages  73--87 of:} Carro, M., \& Pe\~{n}a, R. (eds), {\em 12\xth
  International Symposium on Practical Aspects of Declarative Languages}.
\newblock PADL '10, vol. 5937.
\newblock Springer-Verlag.

\bibitem[\protect\citename{Dieterle {\em et~al.}\relax,
  }2016]{dieterle_horstmeyer_loogen_berthold_2016}
Dieterle, M., Horstmeyer, T., Loogen, R., \& Berthold, J. (2016).
\newblock Skeleton composition versus stable process systems in {Eden}.
\newblock {\em Journal of functional programming}, {\bf 26}.

\bibitem[\protect\citename{Dieterle {\em et~al.}\relax,
  }2010b]{dieterle2010skeleton}
Dieterle, Mischa, Berthold, Jost, \& Loogen, Rita. (2010b).
\newblock {\em A skeleton for distributed work pools in {Eden}}.
\newblock FLOPS '10.
\newblock Springer.
\newblock Pages  337--353.

\bibitem[\protect\citename{Dieterle {\em et~al.}\relax, }2013]{Dieterle2013}
Dieterle, Mischa, Horstmeyer, Thomas, Berthold, Jost, \& Loogen, Rita. (2013).
\newblock {\em Iterating skeletons}.
\newblock IFL '12.
\newblock Springer.
\newblock Pages  18--36.

\bibitem[\protect\citename{Dinan {\em et~al.}\relax,
  }2009]{Dinan:2009:SWS:1654059.1654113}
Dinan, James, Larkins, D.~Brian, Sadayappan, P., Krishnamoorthy, Sriram, \&
  Nieplocha, Jarek. (2009).
\newblock Scalable work stealing.
\newblock {\em Pages  53:1--53:11 of:} {\em Proceedings of the Conference on
  High Performance Computing Networking, Storage and Analysis}.
\newblock SC '09.
\newblock ACM.

\bibitem[\protect\citename{Foltzer {\em et~al.}\relax,
  }2012]{Foltzer:2012:MPC:2398856.2364562}
Foltzer, Adam, Kulkarni, Abhishek, Swords, Rebecca, Sasidharan, Sajith, Jiang,
  Eric, \& Newton, Ryan. (2012).
\newblock A meta-scheduler for the {Par}-monad: Composable scheduling for the
  heterogeneous cloud.
\newblock {\em {SIGPLAN} not.}, {\bf 47}(9), 235--246.

\bibitem[\protect\citename{Geimer {\em et~al.}\relax, }2010]{Geimer2010}
Geimer, M., Wolf, F., Wylie, B. J.~N., {\'A}brah{\'a}m, E., Becker, D., \&
  Mohr, B. (2010).
\newblock The {Scalasca} performance toolset architecture.
\newblock {\em Concurrency and computation: Practice and experience}, {\bf
  22}(6).

\bibitem[\protect\citename{Gentleman, }1978]{Gentleman1978}
Gentleman, W.~Morven. (1978).
\newblock Some complexity results for matrix computations on parallel
  processors.
\newblock {\em Journal of the acm}, {\bf 25}(1), 112--115.

\bibitem[\protect\citename{Gorlatch, }1998]{Gorlatch1998}
Gorlatch, Sergei. (1998).
\newblock Programming with divide-and-conquer skeletons: A case study of {FFT}.
\newblock {\em Journal of supercomputing}, {\bf 12}(1-2), 85--97.

\bibitem[\protect\citename{Gorlatch \& Bischof, }1998]{Gorlatch}
Gorlatch, Sergei, \& Bischof, Holger. (1998).
\newblock A generic {MPI} implementation for a data-parallel skeleton: Formal
  derivation and application to {FFT}.
\newblock {\em Parallel processing letters}, {\bf 8}(4).

\bibitem[\protect\citename{Hammond {\em et~al.}\relax,
  }2003]{doi:10.1142/S0129626403001380}
Hammond, Kevin, Berthold, Jost, \& Loogen, Rita. (2003).
\newblock Automatic skeletons in {Template} {Haskell}.
\newblock {\em Parallel processing letters}, {\bf 13}(03), 413--424.

\bibitem[\protect\citename{Harris {\em et~al.}\relax,
  }2007]{harris2007parallel}
Harris, Mark, Sengupta, Shubhabrata, \& Owens, John~D. (2007).
\newblock Parallel prefix sum (scan) with {CUDA}.
\newblock {\em {GPU} gems}, {\bf 3}(39), 851--876.

\bibitem[\protect\citename{Harris {\em et~al.}\relax,
  }2005]{Harris:2005:CMT:1065944.1065952}
Harris, Tim, Marlow, Simon, {Peyton Jones}, Simon, \& Herlihy, Maurice. (2005).
\newblock Composable memory transactions.
\newblock {\em Pages  48--60 of:} {\em Proceedings of the 10\xth {ACM SIGPLAN}
  Symposium on Principles and Practice of Parallel Programming}.
\newblock PPoPP '05.
\newblock ACM.

\bibitem[\protect\citename{Hey, }1990]{Hey1990185}
Hey, Anthony J.~G. (1990).
\newblock Experiments in {MIMD} parallelism.
\newblock {\em Future generation computer systems}, {\bf 6}(3), 185--196.

\bibitem[\protect\citename{Hippold \& R{\"u}nger, }2006]{Hippold2006}
Hippold, J., \& R{\"u}nger, G. (2006).
\newblock Task pool teams: A hybrid programming environment for irregular
  algorithms on {SMP} clusters.
\newblock {\em Concurrency and computation: Practice and experience}, {\bf 18},
  1575--1594.

\bibitem[\protect\citename{Horstmeyer \& Loogen, }2013]{Horstmeyer2013}
Horstmeyer, Thomas, \& Loogen, Rita. (2013).
\newblock Graph-based communication in {Eden}.
\newblock {\em Higher-order and symbolic computation}, {\bf 26}(1), 3--28.

\bibitem[\protect\citename{Huang {\em et~al.}\relax, }2007]{Huang2007}
Huang, Liwen, Hudak, Paul, \& Peterson, John. (2007).
\newblock {\em {HPorter}: Using arrows to compose parallel processes}.
\newblock Berlin, Heidelberg: Springer Berlin Heidelberg.
\newblock Pages  275--289.

\bibitem[\protect\citename{Hudak {\em et~al.}\relax, }2003]{Hudak2003}
Hudak, Paul, Courtney, Antony, Nilsson, Henrik, \& Peterson, John. (2003).
\newblock {\em Arrows, robots, and functional reactive programming}.
\newblock Springer.
\newblock Pages  159--187.

\bibitem[\protect\citename{Hughes, }2000]{HughesArrows}
Hughes, John. (2000).
\newblock Generalising monads to arrows.
\newblock {\em Science of computer programming}, {\bf 37}(1–3), 67--111.

\bibitem[\protect\citename{Hughes, }2005]{Hughes2005}
Hughes, John. (2005).
\newblock {\em Programming with arrows}.
\newblock AFP '04.
\newblock Springer.
\newblock Pages  73--129.

\bibitem[\protect\citename{Jacobs {\em et~al.}\relax,
  }2009]{jacobs_heunen_hasuo_2009}
Jacobs, Bart, Heunen, Chris, \& Hasuo, Ichiro. (2009).
\newblock Categorical semantics for arrows.
\newblock {\em Journal of functional programming}, {\bf 19}(3-4), 403--438.

\bibitem[\protect\citename{Janjic {\em et~al.}\relax, }2013]{janjic2013space}
Janjic, Vladimir, Brown, Christopher~Mark, Neunhoeffer, Max, Hammond, Kevin,
  Linton, Stephen~Alexander, \& Loidl, Hans-Wolfgang. (2013).
\newblock Space exploration using parallel orbits: a study in parallel symbolic
  computing.
\newblock {\em Parallel computing}.

\bibitem[\protect\citename{Karasawa \& Iwasaki, }2009]{5361825}
Karasawa, Y., \& Iwasaki, H. (2009).
\newblock A parallel skeleton library for multi-core clusters.
\newblock {\em Pages  84--91 of:} {\em International Conference on Parallel
  Processing 2009}.

\bibitem[\protect\citename{Keller {\em et~al.}\relax,
  }2010]{Keller:2010:RSP:1932681.1863582}
Keller, Gabriele, Chakravarty, Manuel~M.T., Leshchinskiy, Roman, {Peyton
  Jones}, Simon, \& Lippmeier, Ben. (2010).
\newblock Regular, shape-polymorphic, parallel arrays in haskell.
\newblock {\em {SIGPLAN} not.}, {\bf 45}(9), 261--272.

\bibitem[\protect\citename{Kuchen, }2002]{Kuchen2002}
Kuchen, Herbert. (2002).
\newblock A skeleton library.
\newblock {\em Pages  620--629 of:} Monien, Burkhard, \& Feldmann, Rainer
  (eds), {\em Parallel Processing}.
\newblock Euro-Par '02.
\newblock Springer.

\bibitem[\protect\citename{Kuper {\em et~al.}\relax,
  }2014]{Kuper:2014:TPE:2666356.2594312}
Kuper, Lindsey, Todd, Aaron, Tobin-Hochstadt, Sam, \& Newton, Ryan~R. (2014).
\newblock Taming the parallel effect zoo: Extensible deterministic parallelism
  with {LVish}.
\newblock {\em {SIGPLAN} not.}, {\bf 49}(6), 2--14.

\bibitem[\protect\citename{Lengauer {\em et~al.}\relax, }1997]{Lengauer1997}
Lengauer, Christian, Gorlatch, Sergei, \& Herrmann, Christoph. (1997).
\newblock The static parallelization of loops and recursions.
\newblock {\em The journal of supercomputing}, {\bf 11}(4), 333--353.

\bibitem[\protect\citename{Li \& Zdancewic, }2006]{1648705}
Li, Peng, \& Zdancewic, S. (2006).
\newblock Encoding information flow in {Haskell}.
\newblock {\em Pages  12--16 of:} {\em 19\xth IEEE Computer Security
  Foundations Workshop}.
\newblock CSFW '06.

\bibitem[\protect\citename{Li \& Zdancewic, }2010]{LI20101974}
Li, Peng, \& Zdancewic, Steve. (2010).
\newblock Arrows for secure information flow.
\newblock {\em Theoretical computer science}, {\bf 411}(19), 1974--1994.

\bibitem[\protect\citename{Lindley {\em et~al.}\relax, }2011]{LINDLEY201197}
Lindley, Sam, Wadler, Philip, \& Yallop, Jeremy. (2011).
\newblock Idioms are oblivious, arrows are meticulous, monads are promiscuous.
\newblock {\em Electronic notes in theoretical computer science}, {\bf 229}(5),
  97--117.

\bibitem[\protect\citename{Linton {\em et~al.}\relax, }2010]{scscp}
Linton, S., Hammond, K., Konovalov, A., Al~Zain, A.~D., Trinder, P., Horn, P.,
  \& Roozemond, D. (2010).
\newblock Easy composition of symbolic computation software: a new lingua
  franca for symbolic computation.
\newblock {\em Pages  339--346 of:} {\em Proceedings of the 2010 International
  Symposium on Symbolic and Algebraic Computation}.
\newblock ISSAC '10.
\newblock ACM Press.

\bibitem[\protect\citename{Liu {\em et~al.}\relax,
  }2009]{Liu:2009:CCA:1631687.1596559}
Liu, Hai, Cheng, Eric, \& Hudak, Paul. (2009).
\newblock Causal commutative arrows and their optimization.
\newblock {\em {SIGPLAN} not.}, {\bf 44}(9), 35--46.

\bibitem[\protect\citename{Lobachev, }2011]{lobachev-phd}
Lobachev, Oleg. (2011).
\newblock {\em Implementation and evaluation of algorithmic skeletons:
  Parallelisation of computer algebra algorithms}.
\newblock Ph.D. thesis, Philipps-Universit{\"a}t Marburg.

\bibitem[\protect\citename{Lobachev, }2012]{Lobachev2012}
Lobachev, Oleg. (2012).
\newblock {\em Parallel computation skeletons with premature termination
  property}.
\newblock FLOPS 2012.
\newblock Springer.
\newblock Pages  197--212.

\bibitem[\protect\citename{Loogen {\em et~al.}\relax,
  }2003]{Eden:SkeletonBookChapter02}
Loogen, R., {Ortega-Mall\'{e}n}, Y., {Pe\~{n}a}, R., Priebe, S., \& Rubio, F.
  (2003).
\newblock {Parallelism Abstractions in Eden}.
\newblock {\em Pages  71--88 of:} Rabhi, F.~A., \& Gorlatch, S. (eds), {\em
  {Patterns and Skeletons for Parallel and Distributed Computing}}.
\newblock Springer-Verlag.

\bibitem[\protect\citename{Loogen, }2012]{Loogen2012}
Loogen, Rita. (2012).
\newblock {\em Eden -- parallel functional programming with {Haskell}}.
\newblock Springer.
\newblock Pages  142--206.

\bibitem[\protect\citename{Loogen {\em et~al.}\relax, }2005]{eden}
Loogen, Rita, Ortega-Mall\'{e}n, Yolanda, \& {Pe\~{n}a-Mar\'{\i}}, Ricardo.
  (2005).
\newblock {Parallel Functional Programming in Eden}.
\newblock {\em {Journal of Functional Programming}}, {\bf 15}(3), 431--475.
\newblock Special Issue on Functional Approaches to High-Performance Parallel
  Programming.

\bibitem[\protect\citename{Lämmel, }2008]{LAMMEL20081}
Lämmel, Ralf. (2008).
\newblock Google’s mapreduce programming model — revisited.
\newblock {\em Science of computer programming}, {\bf 70}(1), 1--30.

\bibitem[\protect\citename{Maier {\em et~al.}\relax,
  }2014]{Maier:2014:HDS:2775050.2633363}
Maier, Patrick, Stewart, Robert, \& Trinder, Phil. (2014).
\newblock The {HdpH} {DSLs} for scalable reliable computation.
\newblock {\em {SIGPLAN} not.}, {\bf 49}(12), 65--76.

\bibitem[\protect\citename{Mainland \& Morrisett,
  }2010]{Mainland:2010:NEC:2088456.1863533}
Mainland, Geoffrey, \& Morrisett, Greg. (2010).
\newblock Nikola: Embedding compiled {GPU} functions in {Haskell}.
\newblock {\em {SIGPLAN} not.}, {\bf 45}(11), 67--78.

\bibitem[\protect\citename{Marlow {\em et~al.}\relax, }2009]{Marlow2009}
Marlow, S., {Peyton Jones}, S., \& Singh, S. (2009).
\newblock Runtime support for multicore {Haskell}.
\newblock {\em {SIGPLAN} not.}, {\bf 44}(9), 65--78.

\bibitem[\protect\citename{Marlow, }2013]{marlow2013parallel}
Marlow, Simon. (2013).
\newblock {\em Parallel and concurrent programming in {Haskell}: Techniques for
  multicore and multithreaded programming}.
\newblock "O'Reilly Media, Inc.".

\bibitem[\protect\citename{Marlow {\em et~al.}\relax, }2011]{par-monad}
Marlow, Simon, Newton, Ryan, \& {Peyton Jones}, Simon. (2011).
\newblock A monad for deterministic parallelism.
\newblock {\em {SIGPLAN} not.}, {\bf 46}(12), 71--82.

\bibitem[\protect\citename{McDonell {\em et~al.}\relax,
  }2015]{McDonell:2015:TRC:2887747.2804313}
McDonell, Trevor~L., Chakravarty, Manuel M.~T., Grover, Vinod, \& Newton,
  Ryan~R. (2015).
\newblock Type-safe runtime code generation: Accelerate to {LLVM}.
\newblock {\em {SIGPLAN} not.}, {\bf 50}(12), 201--212.

\bibitem[\protect\citename{Michael {\em et~al.}\relax,
  }2009]{Michael:2009:IWS:1594835.1504186}
Michael, Maged~M., Vechev, Martin~T., \& Saraswat, Vijay~A. (2009).
\newblock Idempotent work stealing.
\newblock {\em {SIGPLAN} not.}, {\bf 44}(4), 45--54.

\bibitem[\protect\citename{Nilsson {\em et~al.}\relax,
  }2002]{Nilsson:2002:FRP:581690.581695}
Nilsson, Henrik, Courtney, Antony, \& Peterson, John. (2002).
\newblock Functional reactive programming, continued.
\newblock {\em Pages  51--64 of:} {\em Proceedings of the 2002 {ACM} {SIGPLAN}
  Workshop on {Haskell}}.
\newblock Haskell '02.
\newblock New York, NY, USA: ACM.

\bibitem[\protect\citename{Olivier \& Prins, }2008]{4625841}
Olivier, S., \& Prins, J. (2008).
\newblock Scalable dynamic load balancing using {UPC}.
\newblock {\em Pages  123--131 of:} {\em 37\xth International Conference on
  Parallel Processing}.

\bibitem[\protect\citename{Paterson, }2001]{Paterson:2001:NNA:507669.507664}
Paterson, Ross. (2001).
\newblock A new notation for arrows.
\newblock {\em {SIGPLAN} not.}, {\bf 36}(10), 229--240.

\bibitem[\protect\citename{Pe{\~n}a \& Rubio, }2001]{Eden:PPDP01}
Pe{\~n}a, R., \& Rubio, F. (2001).
\newblock {Parallel Functional Programming at Two Levels of Abstraction}.
\newblock {\em Pages  187--198 of:} {\em PPDP'01 --- Intl.\ Conf.\ on
  Principles and Practice of Declarative Programming}.

>>>>

% More experiences with seamless porting of parallel PArrows-based programs across the backends are welcome.
% Of course, we are ourselves working on expanding
In ongoing work we are expanding 
both our skeleton library and the number of skeleton-based parallel programs that use our DSL.
% to be portable across flavours of parallel Haskells.
It would also be interesting to see a hybrid of PArrows and Accelerate \cite{McDonell:2015:TRC:2887747.2804313}.
Ports of our approach to other languages such as Frege, Eta, or Java directly are at an early development stage.
	%\pagebreak
        \bibliographystyle{jfp}
	\bibliography{references,main}
        \appendix
	\section{Utility Arrows}\label{utilfns}
Following are definitions of some utility Arrows used in this paper that have been left out for brevity.
We start with the \ensuremath{\Varid{second}} combinator from \citet{HughesArrows}, which is a mirrored version of \ensuremath{\Varid{first}}, which is for example used in the definition of \ensuremath{\mathbin{*\!*\!*}}: 
\begin{hscode}\SaveRestoreHook
\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
\column{9}{@{}>{\hspre}l<{\hspost}@{}}%
\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
\>[B]{}\Varid{second}\mathbin{::}\Conid{Arrow}\;\Varid{arr}\Rightarrow \Varid{arr}\;\Varid{a}\;\Varid{b}\to \Varid{arr}\;(\Varid{c},\Varid{a})\;(\Varid{c},\Varid{b}){}\<[E]%
\\
\>[B]{}\Varid{second}\;\Varid{f}\mathrel{=}\Varid{arr}\;\Varid{swap}\mathbin{>\!\!>\!\!>}\Varid{first}\;\Varid{f}\mathbin{>\!\!>\!\!>}\Varid{arr}\;\Varid{swap}{}\<[E]%
\\
\>[B]{}\hsindent{9}{}\<[9]%
\>[9]{}\mathbf{where}\;\Varid{swap}\;(\Varid{x},\Varid{y})\mathrel{=}(\Varid{y},\Varid{x}){}\<[E]%
\ColumnHook
\end{hscode}\resethooks
